---
title: "ResIN-VIGNETTE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ResIN-VIGNETTE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
if(!require("pacman")) install.packages('pacman')
library(pacman)
p_load(tidyverse)
p_load(devtools)


install_github("pwarncke77/ResIN")


```

Welcome to the ResIN R-Package vignette. Here we'll showcase and comment on some of the features available in the ResIN package.

## A simple example with simmualted data: 

### Let's first simmulate a slightly noisy dataset of 8 Likert-scale items for 1000 hypothetical respondents:

```{r simmulate data, echo=FALSE, message=FALSE, warning=FALSE}
 set.seed(42)
 n <- 1000
 k <- 8

 latent_dgp <- rnorm(1000, 0, 1)

 ## Data for k manifest indicators measured with some error
 cont_data <- matrix(rep(latent_dgp, k), n, k)
 cont_data <- apply(cont_data, 2, function(x) {
   x + rnorm(1000, 0, 0.2)}) ## 20% error

 ## Ordinal, 5-point Likert-scale items
 lik_maker <- function(x) {
   sort(runif(x, min = -2, max = 2))}

 lik_data <- matrix(NA, n, k)

 for(i in 1:ncol(lik_data)) {
   lik_data[, i] <- findInterval(cont_data[, i], vec = c(-Inf, lik_maker(4), Inf))
 }


```

### Running the ResIN algorithm

The first argument supplied to the \code{ResIN} function is simply the data-frame containing all (survey) items we would like to include in the analysis. Here, we also chose the correlation method that calculates the item-pair associations and specified that we'd like to estimate network descriptive statistics. All this can be done in one line of code.

```{r run ResIN, echo=FALSE, message=FALSE, warning=FALSE}
 # Apply the ResIN function to toy Likert data:

 output <- ResIN(lik_data, cor_method = "spearman", network_stats = TRUE)



```

### Visualization

We can now freely investigate the ResIN output. Let's make use of the convenient, ggplot-ready edge-list stored in the \code{ResIN} output to visualize the result. Because we want to take advantage of our network statistics, we first re-order the ggplot dataframe 

```{r generate ggplots, echo=FALSE, message=FALSE, warning=FALSE}

# Create a basic outcome plot with ggplot
 output$ggplot_frame <- output$ggplot_frame[order(output$ggplot_frame$Strength,
                                                  decreasing = FALSE), ]
 ResIN_plot <- ggplot(output$ggplot_frame) +
   geom_curve(data = output$ggplot_frame, aes(x = from.x, xend = to.x, y = from.y,
                                              yend = to.y, linewidth = weight,
                                              color = Strength), curvature = 0.2) +
   geom_point(aes(x = from.x, y = from.y, shape = as.factor(cluster)), size = 8)+
   geom_point(aes(x = to.x, y = to.y), size = 8)+
   geom_text(data = output$ggplot_frame, aes(x = from.x, y = from.y, label = from),
             size = 3, color = "white") +
   geom_text(data = output$ggplot_frame, aes(x = to.x, y = to.y, label = to),
             size = 3, color = "white") +
   ggtitle("ResIN example  plot")+
   theme_dark()+
   theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
         axis.text.y = element_blank(), axis.title.y = element_blank(),
         axis.ticks = element_blank(), panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(), legend.position = "none",
         legend.text = element_blank(), plot.title = element_text(hjust = 0.5))

 ResIN_plot
 
 ```
 
 
